Visual diagnostics for constrained optimisation with application to guided tours

### 2020-10-09 paper/paper.Rmd
remove codes, comments, list of plot elements that are unncessary, kol_1d_2var_beter_sim2 data

[Are we using  projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below? Or we are focusing on the optimisation problem only within the projection pursuit context? Some of the problems listed below are also applicable to optimization problem in general too. ppp]

@@

```{r tour-path, out.width="60%", out.height="40%", fig.align='center', fig.cap="\\label{tour-path}An illustration of the tour path"}
knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.png"),)
```

@@

[outside the paper: I find the notation of current/target basis is confusing because the target basis in round $j$ becomes the current basis in round $j+1$. Also, when we start to have polish, the target basis may not be the current basis in the next round... The place where current/target is most appropriate is probably when describing the interpolation where the first one is always the current basis and the last is always the target basis. I think it is better to leave this language in the code]

@@

Another option is to use summarisation for each iteration. Boxplot is a suitable candidate that provides five points summary of the data, however, there are two pieces of information missing from the boxplot: 1) It does not report the number of points, and 2) the position of the last basis. This could be remedied by adding more layers using the concept of grammar of graphics. A label geometry is added at the bottom of the plot to show the number of points in each iteration and a line geometry links the last basis in each iteration. Further, an option to switch between displaying points and boxplot geometry is helpful since point geometry can be more intuitive for the iteration with few observations. This is achieved via a `cutoff` parameter. Below we define the searching point plot based on four different component layers

- Layer 1: boxplot geom
  - data: group by $j$ and filter the observations in the groups that have count greater than `cutoff = 15`.
  - x: $j$ is mapped to the x-axis
  - y: the statistical transformed index value: $Q_{I^{\prime}_t}(q)$ is mapped to the y-axis where $Q_X(q)$, $q = 0, 25, 50, 75, 100$ finds the qth-quantile of $X$ and $I^{\prime}_t$ denotes the index value of all the searching bases defined in Matrix \ref{eq:data-structure}.

- Layer 2: point geom
  - data: group by $j$ and filter the observations in the group that have count less than `cutoff = 15`.
  - x: $j$ is mapped to the x-axis
  - y: $I$ is mapped to the y-axis

- Layer 3: line geom
  - data: filter the points with the highest index value in group $j$
  - x: $j$ is mapped to the x-axis
  - y: $I$ is mapped to the y-axis

- Layer 4: label geom
  - data: A count table of the number of observation in each iteration
  - x: $j$ is mapped to the x-axis
  - y: the y-axis for each label is $0.99 * \text{MIN}$, where $\text{MIN}$ is the smallest index value in the whole data object
  - label: the number of observation is mapped to the label

<!-- *Example: exploring searching points* We choose variables `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour. Parameter `search_f = tour::search_better` and `max.tries = 500` is used. The index value of the searching points are shown in Figure \ref{toy-search}. The label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`). We learn that the `search_better` quickly finds better projection basis with higher index value at first and then takes longer to find a better one later. -->

@@

```{r toy-search, fig.height=5.5, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs. In the upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read."}

dt <- holes_2d_better_max_tries %>%
  filter(info != "interpolation") %>% 
  mutate(id = row_number())

p1 <- dt %>% explore_trace_search() + 
  ggtitle("satisfactory") + 
  scale_color_botanical(palette = "daisy")

p2 <- holes_2d_better %>% explore_trace_search() + 
  ggtitle("not satisfactory") + 
  scale_color_botanical(palette = "daisy")

p1 / p2

@@

Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation. The plot definition is as follows:

- Layer 1: point geom
  - data: filter the observations with $S = \text{interpolation}$ and mutate $t$ to be the row number of the subsetted tibble
  - x: $t$ is mapped to the x-axis
  - y: $I$ is mapped to the y-axis

- Layer 2: line geom
  - using line geometry for the same data and aesthetics

Figure \ref{toy-interp} presents the interpolation of three different tour paths. The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.  While in the middle plot, the increases towards the target basis is not monotonical in the last two iterations and interpolated basis with higher index value can be found on the tour path. The lower plot is constructed using `search_better_random`, where a basis with smaller index value  has a probabilistic chance of being accepted and we can observe a much more involved pattern. In iteration three, the probabilistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probabilistic acceptance, the interpolation is now non-monotonical. When the target basis has a higher index value, iteration eight reaches this basis by first decreasing the index value. While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnosis of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.

```{r toy-interp, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
p1 <- holes_2d_better_max_tries %>% explore_trace_interp() + theme(legend.position = "none") + 
  scale_color_botanical(palette = "fern")

p2 <- interrupt_no %>% explore_trace_interp() + theme(legend.position = "none") + 
  scale_color_botanical(palette = "fern")

p3 <- holes_2d_better_random %>% explore_trace_interp() + theme(legend.position = "none") + 
  scale_color_botanical(palette = "fern")

p1 / p2 / p3
```

\

@@

The visualisation can thus be defined as

@@

- Layer 1: point geom
  - data: subset the basis of interest and arrange into a matrix format; perform PCA on the basis matrix and compute the projected basis on the first two principal components; bind the variables from the original global object and form a tibble
  - x: the projected basis on the first principal component
  - y: the projected basis on the second principal component
  - colour: $V$ is mapped to the colour aesthetic

@@

```{r toy-pca, fig.height = 4, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
# path1 <- holes_1d_geo$basis %>% 
#   flatten_dbl() %>% 
#   matrix(ncol = 5, byrow = TRUE)
# path2 <- holes_1d_better$basis %>% 
#   flatten_dbl() %>% 
#   matrix(ncol = 5, byrow = TRUE)
# set.seed(1)
# sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
# theoretical <- matrix(c(0, 1, 0, 0, 0,
#                         0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
# 
# dt <- rbind(sphere, path1, path2, theoretical)
# pca <- dt %>% prcomp(scale. = TRUE)
# #summary(pca)
# pca_pred <- predict(pca) %>%
#   as_tibble() %>%
#   mutate(method = c(rep("sphere", nrow(sphere)),
#                     holes_1d_geo$method,
#                     holes_1d_better$method,
#                     rep("theoretical", 2))) %>%
#   group_by(method) %>%
#   mutate(id = row_number(),
#          id = ifelse(method == "sphere", 1, id)) %>%
#   ungroup()
# 
# p <- pca_pred %>%
#   ggplot(aes(x = PC1, y = PC2, color = method))+
#   geom_point() +
#   geom_point(data = pca_pred %>% filter(method == "theoretical"), 
#              size = 10) + 
#   geom_point(data = pca_pred %>% filter(method %in% c("search_better", "search_geodesic") & id == 1), size = 10) + 
#   theme(aspect.ratio = 1) + 
#   scale_color_botanical(palette = "cherry")
# 
# p

@@

Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.

<!-- all the plots above can be animated if wanted - importance is how the data is moving around  -->

<!-- show an example of some snap shots  of the animation and provide a link... for animated pca plot -->

<!-- This is particularly useful when plotting the PCA plot since it allways us to tract the progress of the search. -->

<!-- The projection bases live in a high dimensional sphere, whch cna be simulated using `geozoo` package. -->

<!-- The reason to simulate random points on the hihg dimensiaon sphere is to emphasize  taht the projection bases lives on the surface of the sphere. -->

```{r fig.cap = "\\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot. With animation, it is easier to track the progrssion from the start to finish in each algorithm.", fig.height=4, fig.align="center"}
# ani <- p + theme(legend.position = "none") +
#   transition_states(id) +
#   shadow_mark()
# 
# ani <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
#                    index = tourr::holes(), raw_data = data) %>% 
#   bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
#                    index = tourr::holes(), raw_data  = data)%>%
#   explore_space_pca(col = method, animate = TRUE)  + 
#   scale_color_botanical(palette = "cherry") 
# 
# animate(ani, nframes = 350, device = "png",
#         renderer = file_renderer("figures/pca/",
#                                  prefix = "pca", overwrite = TRUE))

@@

#animate_xy(dt, col = color, cex = cex, tour_path = grand_tour()) # trial

@@

<!-- In the previous two sections, only the iterator and the index value are mapped onto the x and y aesthetics of the plot; while more aesthetics i.e. colour, could be added to compare other parameters in the global object. -->

<!-- Two examples are shown below to explore and compare different searching methods and neighbourhood parameter alpha.  -->

Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood. This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough. The following code demonstrates the usage of `search_polish`:

```{r sample-code, eval = FALSE, echo = TRUE}
library(tourr)
library(ferrn)

set.seed(123456)
holes_2d_geo <- 
  animate_xy(boa6,
             tour_path = guided_tour(holes(), d = 2, 
                                     search_f = search_geodesic),
             rescale = FALSE, verbose = TRUE)

last_basis <- get_best(holes_2d_geo) %>% pull(basis) %>% .[[1]]

set.seed(123456)
holes_2d_geo_polish <- 
  animate_xy(boa6, 
             tour_path = guided_tour(holes(), d = 2, 
                                     search_f = search_polish),
             rescale = FALSE, verbose = TRUE, start = last_basis)
```

@@

```{r kol-cdf,fig.asp=0.5, fig.cap = "\\label{kol-cdf}Trace plot of search_geodesic and search_better on 1D projection problem for a noisy index: norm_kol. Search_geodesic has varely made progression while search_better has successfully found the informative "}
# set.seed(123456)
# kol_1d_geo <- 
#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
#                                              search_f =  search_geodesic, max.tries = 100),
#                rescale = FALSE, verbose = TRUE)
# 
# set.seed(123456)
# kol_1d_better <- 
#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
#                                              search_f =  search_better, max.tries = 100),
#                rescale = FALSE, verbose = TRUE)
# save(kol_1d_geo, file = here::here("data", "kol_1d_geo.rda"))
# save(kol_1d_better, file = here::here("data", "kol_1d_better.rda"))

@@

```{r fig.height= 8}
set.seed(123)
seed <- sample.int(10000, 20)
# kol_1d_2var_better_random_sim <- list()
# kol_1d_2var_better_random_sim_polish <- list()
# for (i in 1: length(seed)){
#   cat("i = ", i, "seed = ", seed[i], "\n")
#   set.seed(seed[i])
#   kol_1d_2var_better_random_sim[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_better, max.tries = 500, t0 = 1e-6, alpha = 0.7),
#                  rescale = FALSE, verbose = TRUE)
# 
# 
#   last_basis <- get_best(kol_1d_2var_better_random_sim[[i]])$basis %>% .[[1]]
# 
#   set.seed(seed[i])
#   kol_1d_2var_better_random_sim_polish[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_polish, max.tries = 200),
#                  rescale = FALSE, verbose = TRUE,
#                  start = last_basis)
#   
# }

@@

```{r kol-pca, fig.height = 5, fig.cap = "\\label{kol-pca}Comparison between search_better_random and two different search_better. In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases. serach_better_random and one of the serach_better have found the global maximum but the other fails. The lower panel shows the PCA plot for each of the three algorithms. The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum. The paths of two search_better are relatively short and simple while the search_better_random has a more sophisticated path. This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance."}

# set.seed(123456)
# kol_1d_2var_better <- 
#   animate_dist(boa6, 
#                tour_path = guided_tour(norm_kol(), d = 1,
#                                        search_f = search_better, max.tries = 100),
#                rescale = FALSE, verbose = TRUE)
# 
# 
# set.seed(1234567)
# kol_1d_2var_better_2 <- 
#   animate_dist(boa6, 
#                tour_path = guided_tour(norm_kol(), d = 1,
#                                        search_f = search_better, max.tries = 100),
#                rescale = FALSE, verbose = TRUE)
# 
# set.seed(12345678)
# kol_1d_2var_better_2 <- 
#   animate_dist(boa6, 
#                tour_path = guided_tour(norm_kol(), d = 1,
#                                        search_f = search_better, max.tries = 100),
#                rescale = FALSE, verbose = TRUE)
# 
# 
# 
# 
# set.seed(123456)
# kol_1d_2var_better_random <- 
#   animate_dist(boa6, 
#                tour_path = guided_tour(norm_kol(), d = 1,
#                                        search_f = search_better_random,
#                                        max.tries = 2000, t0 = 1e-5),
#                rescale = FALSE, verbose = TRUE)
# 
# save(kol_1d_2var_better, file = here::here("data", "kol_1d_2var_better.rda"))
# save(kol_1d_2var_better_2, file = here::here("data", "kol_1d_2var_better_2.rda"))
# save(kol_1d_2var_better_random, file = here::here("data", "kol_1d_2var_better_random.rda"))
# 
# p1 <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
#           kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
#           kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
#   explore_trace_interp(group = algo, col = algo) + 
#   facet_wrap(vars(algo), ncol = 3)  + 
#   scale_color_botanical(palette = "banksia")  + 
#   theme(legend.position = "none", aspect.ratio = 1)
# 
# pca <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
#           kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
#           kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
#   bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
#                                         index = tourr::norm_kol(),
#                                         raw_data = boa6) %>% 
#    bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
#                                         index = tourr::norm_kol(),
#                                         raw_data = boa6) %>% 
#   compute_pca()

dt <- bind_rows(all_tidy_better %>% filter(seed == 1842) %>% mutate(algo = "better V2"),
          all_tidy_better %>% filter(seed == 9982) %>% mutate(algo = "better V7"),
          all_tidy %>% filter(seed == 4761) %>% mutate(algo = "better random"))

index <- tourr::norm_kol() 
theo1 <- index(as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6))
theo2 <- index(as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6))
p1 <- dt %>% 
  explore_trace_interp(group = algo, col = algo) + 
  facet_wrap(vars(algo), ncol = 3)  + 
  geom_hline(yintercept = theo1, alpha = 0.5, linetype = 2) + 
  geom_hline(yintercept = theo2, alpha = 0.5, linetype = 2) + 
  scale_color_botanical(palette = "banksia")  + 
  theme(legend.position = "none", aspect.ratio = 1)

pca <- dt %>% 
  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
                                        index = tourr::norm_kol(),
                                        raw_data = boa6) %>% 
   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
                                        index = tourr::norm_kol(),
                                        raw_data = boa6) %>% 
  compute_pca()
random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-algo)
theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-algo)

p2 <- pca$aug %>% filter(method != "randomly_generated" & info != "theoretical") %>% 
  ggplot(aes(x = PC1, y = PC2, col = algo)) + 
  geom_point(data = random, col = "grey") +
  geom_point(aes(col = algo), alpha = 0.05) + 
  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
  geom_point(data = pca$aug %>% get_best(group = algo), size = 5) +
  geom_point(data = pca$aug %>% get_start() %>% filter(method %in% c("search_better_random", "search_better")),  size = 3) +
  facet_wrap(vars(algo)) +
  theme_void() + 
  theme(legend.position = "none", aspect.ratio = 1) + 
  scale_color_botanical(palette = "banksia")

wrap_plots(p1, p2, ncol = 1)

@@

(further notice on the implementation in the tour package will be discussed in the tour developer meeting.)
The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing `tourr` package. The diagnostics of the algorithm via plots and other auxiliary functions have been implemented in a new pacakge, `ferrn`.

@@

```{r glb-obj-1}
holes_1d_better %>% 
  dplyr::select(id, basis, index_val, info, tries, loop, method, alpha) %>% 
  head(10) 
```

@@

  <!-- This binding is used in the space plots to compare the position of the theoretical best basis to the ones being searched by the algorithm. The following chunk presents its usage -->
<!-- These are also useful in the space plots since it preserves the spherical structure of the spaces of the projection bases, in the computation of PCA for `explore_space_pca()` and animation of `explore_space_tour()`.  -->

@@

```{r eval = FALSE}
# Below are codes that generate the plots in Section 4 
# using functions from ferrn package
bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
          holes_2d_better %>% mutate(max_tries = 25)) %>% 
  explore_trace_search(group = max_tries)
```

```{r eval = FALSE}
bind_rows(holes_2d_better_max_tries %>% mutate(group = "max_tries: 500"), 
          interrupt_no %>% mutate(group = "no interruption"), 
          holes_2d_better_random %>% mutate(group = "better_random")) %>% 
  mutate(group = fct_relevel(group, c("max_tries: 500", 
                                      "no interruption", 
                                      "better_random"))) %>% 
  explore_trace_interp(group = group) + 
  theme(legend.position = "none")
```

```{r eval = FALSE}
bind_rows(holes_1d_geo, holes_1d_better) %>% 
  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
                   index = tourr::holes(), raw_data = data) %>% 
  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
                   index = tourr::holes(), raw_data  = data)%>%
  explore_space_pca(col = method)  + 
  scale_color_botanical(palette = "cherry")
```

```{r eval = FALSE}
bind_rows(holes_1d_geo, holes_1d_better) %>% 
  explore_space_tour(color = method)
```

>>>>

@@

```{r tour-path, out.height="30%", out.width="60%", fig.align='center', fig.cap="\\label{tour-path}An illustration of the tour path"}
knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.png"))
```

@@

@@

Another option is to use summarisation for each iteration. Boxplot is a suitable candidate that provides five points summary of the data, however, there are two pieces of information missing from the boxplot: 1) It does not report the number of points, and 2) the position of the last basis. This could be remedied by adding more layers using the concept of grammar of graphics. A label geometry is added at the bottom of the plot to show the number of points in each iteration and a line geometry links the last basis in each iteration. Further, an option to switch between displaying points and boxplot geometry is helpful since point geometry can be more intuitive for the iteration with few observations. This is achieved via a `cutoff` parameter.

@@

```{r toy-search, fig.height=3, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs. In the upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read."}
bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
          holes_2d_better %>% mutate(max_tries = 25)) %>% 
  explore_trace_search(group = max_tries) + 
  scale_color_botanical(palette = "daisy") +
  facet_wrap(vars(max_tries), ncol = 2)
```

@@

Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation. Figure \ref{toy-interp} presents the interpolation of three different tour paths. The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.  While in the middle plot, the increases towards the target basis is not monotonical in the last two iterations and interpolated basis with higher index value can be found on the tour path. The lower plot is constructed using `search_better_random`, where a basis with smaller index value  has a probabilistic chance of being accepted and we can observe a much more involved pattern. In iteration three, the probabilistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probabilistic acceptance, the interpolation is now non-monotonical. When the target basis has a higher index value, iteration eight reaches this basis by first decreasing the index value. While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnosis of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.

```{r toy-interp, fig.height = 3, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
bind_rows(holes_2d_better_max_tries %>% mutate(group = "max_tries: 500"), 
          interrupt_no %>% mutate(group = "no interruption"), 
          holes_2d_better_random %>% mutate(group = "better_random")) %>% 
  mutate(group = fct_relevel(group, c("max_tries: 500", 
                                      "no interruption", 
                                      "better_random"))) %>% 
  explore_trace_interp(group = group) + 
  theme(legend.position = "none") + 
  scale_color_botanical(palette = "fern") + 
  facet_wrap(vars(group), ncol = 3)
```

@@

@@

@@

```{r toy-pca, fig.height = 3, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
# path1 <- holes_1d_geo$basis %>% 
#   flatten_dbl() %>% 
#   matrix(ncol = 5, byrow = TRUE)
# path2 <- holes_1d_better$basis %>% 
#   flatten_dbl() %>% 
#   matrix(ncol = 5, byrow = TRUE)
# set.seed(1)
# sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
# theoretical <- matrix(c(0, 1, 0, 0, 0,
#                         0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
# 
# dt <- rbind(sphere, path1, path2, theoretical)
# pca <- dt %>% prcomp(scale. = TRUE)
# #summary(pca)
# pca_pred <- predict(pca) %>%
#   as_tibble() %>%
#   mutate(method = c(rep("sphere", nrow(sphere)),
#                     holes_1d_geo$method,
#                     holes_1d_better$method,
#                     rep("theoretical", 2))) %>%
#   group_by(method) %>%
#   mutate(id = row_number(),
#          id = ifelse(method == "sphere", 1, id)) %>%
#   ungroup()
# 
# p <- pca_pred %>%
#   ggplot(aes(x = PC1, y = PC2, color = method))+
#   geom_point() +
#   geom_point(data = pca_pred %>% filter(method == "theoretical"), 
#              size = 10) + 
#   geom_point(data = pca_pred %>% filter(method %in% c("search_better", "search_geodesic") & id == 1), size = 10) + 
#   theme(aspect.ratio = 1) + 
#   scale_color_botanical(palette = "cherry")
# 
# p

@@

Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order. In Figure \ref{toy-pca-animated}, ...

```{r fig.height = 3, fig.cap = "\\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot. With animation, it is easier to track the progrssion from the start to finish in each algorithm."}
# ani <- p + theme(legend.position = "none") +
#   transition_states(id) +
#   shadow_mark()
# 
# ani <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
#                    index = tourr::holes(), raw_data = data) %>% 
#   bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
#                    index = tourr::holes(), raw_data  = data)%>%
#   explore_space_pca(col = method, animate = TRUE)  + 
#   scale_color_botanical(palette = "cherry") 
# 
# animate(ani, nframes = 350, device = "png",
#         renderer = file_renderer("figures/pca/",
#                                  prefix = "pca", overwrite = TRUE))

@@

# 
# bind_rows(holes_1d_geo, holes_1d_better) %>% 
#   explore_space_tour(color = method)
#animate_xy(dt, col = color, cex = cex, tour_path = grand_tour()) # trial

@@

Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood. This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough.

@@

```{r kol-cdf,fig.asp=0.5, fig.height = 3,  fig.cap = "\\label{kol-cdf}Trace plot of search_geodesic and search_better on 1D projection problem for a noisy index: norm_kol. Search_geodesic has varely made progression while search_better has successfully found the informative "}
# set.seed(123456)
# kol_1d_geo <- 
#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
#                                              search_f =  search_geodesic, max.tries = 100),
#                rescale = FALSE, verbose = TRUE)
# 
# set.seed(123456)
# kol_1d_better <- 
#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
#                                              search_f =  search_better, max.tries = 100),
#                rescale = FALSE, verbose = TRUE)
# save(kol_1d_geo, file = here::here("data", "kol_1d_geo.rda"))
# save(kol_1d_better, file = here::here("data", "kol_1d_better.rda"))

@@

```{r fig.height = 8}
set.seed(123)
seed <- sample.int(10000, 20)
# kol_1d_2var_better_random_sim2 <- list()
# kol_1d_2var_better_random_sim_polish2 <- list()
# for (i in 1: length(seed)){
#   cat("i = ", i, "seed = ", seed[i], "\n")
#   set.seed(seed[i])
#   kol_1d_2var_better_random_sim2[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_better_random, max.tries = 200),
#                  rescale = FALSE, verbose = TRUE)
# 
# 
#   last_basis <- get_best(kol_1d_2var_better_random_sim2[[i]])$basis %>% .[[1]]
# 
#   set.seed(seed[i])
#   kol_1d_2var_better_random_sim_polish2[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_polish, max.tries = 200),
#                  rescale = FALSE, verbose = TRUE,
#                  start = last_basis)
# 
# }
# save(kol_1d_2var_better_random_sim2, file = here::here("data", "kol_1d_2var_better_sim2.rda"))
# save(kol_1d_2var_better_random_sim_polish2, file = here::here("data", "kol_1d_2var_better_sim_polish2.rda"))

polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish2,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
better_list <- purrr::map2(kol_1d_2var_better_random_sim2, seed,  ~.x %>% dplyr ::mutate(seed = .y)) 
polish_tidy <- do.call(dplyr::bind_rows, polish_list)
better_tidy <- do.call(dplyr::bind_rows, better_list)
all_tidy <- bind_rows(better_tidy, polish_tidy)

best <- get_best(all_tidy, group = seed) %>% arrange(index_val)

```{r fig.height= 8}
set.seed(123)
seed <- sample.int(10000, 20)
# kol_1d_2var_better_random_sim <- list()
# kol_1d_2var_better_random_sim_polish <- list()
# for (i in 1: length(seed)){
#   cat("i = ", i, "seed = ", seed[i], "\n")
#   set.seed(seed[i])
#   kol_1d_2var_better_random_sim[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_better_random, max.tries = 500, t0 = 1e-6, alpha = 0.7),
#                  rescale = FALSE, verbose = TRUE)
# 
# 
#   last_basis <- get_best(kol_1d_2var_better_random_sim[[i]])$basis %>% .[[1]]
# 
#   set.seed(seed[i])
#   kol_1d_2var_better_random_sim_polish[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_polish, max.tries = 200),
#                  rescale = FALSE, verbose = TRUE,
#                  start = last_basis)
#   
# }

@@

Comparison between search_better_random and two different search_better. In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases. serach_better_random and one of the serach_better have found the global maximum but the other fails. The lower panel shows the PCA plot for each of the three algorithms. The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum. The paths of two search_better are relatively short and simple while the search_better_random has a more sophisticated path. This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance

@@

The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing `tourr` package. The diagnostics of the algorithm via plots and other auxiliary functions have been implemented in a new pacakge, `ferrn`.

@@

```{r glb-obj-1}
holes_1d_better %>% 
  dplyr::select(id, basis, index_val, info, tries, loop, method, alpha) %>% 
  head(5) 
```

@@

@@
